---
title: "Lab 1"
author: "Haley Jeppson, Ashish Jain, Vinny Paris"
date: "January 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


***


## Question 1: Building a Book


#### VINNY's problems:

While trying to compile H. Wickham's Advanced R book I run into the following issue:

`grep: Inconsolata.log: No such file or directory
mktextfm: mf-nowin -progname=mf \mode:=ljfour; mag:=1; nonstopmode;  `
`input Inconsolata' failed to make Inconsolata.tfm.`

I am using a mac and I believe that the problem is a font does not naturally exist on my computer. Certain pieces of the book were created though (including the dedictaion and copyright section). I have downloaded the Inconsolata file but am stuck figuring out how to move it to the Font directory in my library. So my question is two fold: 1) Am I on the right track? 2) How do I move a file around on a mac? Thank you!


#### Ashish's Problems and Solutions:

I also faced the same problem while knitting the H. Wickham's Advanced R book. The book is using the "Inconsolata" font which is not a default font in mac. So in order to debug it, I googled it.

**"Whenever in trouble use GOOGLE"**

While searching through the internet we found the solution. First, we downloaded the Inconsolota font from [link](https://cooltext.com/Download-Font-Inconsolata). After that, we followed [these](http://superuser.com/questions/56787/how-to-install-and-find-fonts-for-mac-os-x) instructions and copied the font file into the following folders.

* /Library/Fonts/
* /Users/username/Library/Fonts

Thank you Haley and Vinny for your help!!

#### Haley's Problems and Solutions:

I encountered the same issues as Vinny and Ashish did when attempting to knit Advanced R. Ashish was able to troubleshoot and realize that we were not able to get the book to knit because we were all missing the "Inconsolata" font. I was fortunate to not be a mac user at this moment. In order to install the font on a Ubuntu platform, I simply had to run the code `sudo apt-get install fonts-inconsolata` in the terminal (thanks stackoverflow!). After doing so, I was able to compile the pdf and discover that the last word on page 21 is "May".


***


## Question 2: Weather Stations

#### Retrieving the Data

The function `tmax` takes the argument `stationid` in the form of six digits, reads in the corresponding file, and cleans the tibble before returning it.


```{r}
library(tidyverse)

tmax <- function(stationid){
  # paste together file name
  station <- paste("https://raw.githubusercontent.com/haleyjeppson/Data-Technologies/master/Lab1/ushcn.v2.5.5.20170118/USH00", stationid, ".raw.tmax", sep="")
  # read in the file
  temps <- read_fwf(station, 
                   col_positions=  fwf_positions(
                     start=c( 1,13, rep(16+9*0:11, each=4) + c(1,7,8,9)), 
                     end  =c(11,16, rep(16+9*0:11, each=4) + c(6,7,8,9))))
  
  # rename the column names
names(temps) <- c("Station", "Year", 
                 paste0(rep(c("Value","DMflag", "QCflag", "DSflag"), 12), 
                        rep(1:12, each=4)))

# Now to tidy the data a little
# First, gather all columns that start with "Value" and create a new column for Month based on number that follows "value"
temps <- temps %>% 
  gather(key="Month", value="Temp_Max", starts_with("Value"))
temps$Month <- as.numeric(gsub("Value", "", temps$Month))
# replace the -9999 temperature values with NAs
temps$Temp_Max <- replace(temps$Temp_Max, 
                          temps$Temp_Max == -9999, NA)
# Reformat temps and create a variable for year and month.
temps$Temp_Max <- temps$Temp_Max/100
temps$YearMonth <- temps$Year+temps$Month/12

# Return cleaned tibble
temps
}

data <- tmax("041758")

head(data)

```



#### Visualizing the temperature records over time

The plots below display data from station 041758 in Chula Vista, California.

```{r temps1}

data %>% ggplot(aes(x = Month, Temp_Max, color= Year)) + 
  geom_path(aes(group= Year), size=.7, alpha=.6) +
  scale_color_gradient(low="steelblue1", high="firebrick1") +
  labs(y="Max temp (Celsius)", title = "Monthly max temperatures") 
```



Description:

```{r temps2}
data %>%  group_by(Year) %>%
  mutate(MeanMaxTemp = mean(Temp_Max)) %>%
  ggplot(aes(x=Year, MeanMaxTemp)) + 
  geom_point(aes(color=MeanMaxTemp)) +
  labs(y="Mean max temp (Celsius)", title = "Yearly mean max temperatures") +
  scale_color_gradient(low="steelblue1", high="firebrick1")

```


Description:

